Is the direction of Granger causal influence the same as the direction of information flow?

### 2015-07-07 paper.tex
Correct spellings.

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Is the direction of Granger causal influence the same as the direction of information flow?}
\author{Praveen Venkatesh\\vpraveen@cmu.edu\\Carnegie Mellon University}
\author{Pulkit Grover\\pulkit@cmu.edu\\Carnegie Mellon University}

@@

In order to determine the direction of causal influence, the Granger Causality metric (ratio of residual variances) from $\{X\}$ to $\{Y\}$ is \emph{compared} to that from $\{Y\}$ to $\{X\}$. The direction of causal influence is taken to be the direction with a greater Granger Causality metric. Further, this direction of causal influence is interpreted to be the direction of informationf flow -- the interpretation we question in this paper. We also note that many of these works use a spectral version of Granger Causality, that supplies this metric as a function of frequency. It then becomes possible to also determine the brain wave frequency at which these influences occur. However, we restrict our analysis to the simpler non-spectral version of Granger Causality, since it is sufficient to for the purpose of our arguments.

@@

\begin{enumerate}
	\item First, Granger Causality suffers from what we call the ``hidden node problem''. If two observed nodes receive causal influences from a third, latent node, then a causal influence may be detected between the observed nodes, even if they are independent of each other~\cite{pearl2009causality-hidden-node}. All nodes need to be observed, therefore, to avoid finding spurious influences.
	\item Second, if the measurements from each node are differentially affected by noise, then the predicted direction of causal influence might be opposite to the true direction (\cite{nalatore2007mitigating,andersson2005testing}). Measurements need to be relatively noiseless and precise in order to obtain the correct direction of causal influece.
	\item Third, subsampling the processes can produce misleading Granger Causal relations~\cite{gong2015discovering}. Performing Granger Causal analysis on subsampled time series can lead one to miss the causal influence. If pre-processing involves subsampling, then this should be done with care.
\end{enumerate}

@@

The Granger Causality and Directed Inforamtion metrics are a function of the regression coefficients estimated from the data by fitting the models given by equations~\ref{eq:ar-model-x-x} and~\ref{eq:ar-model-x-x-theta}. In our analysis, we described the data-generation model: the algorithm inspired by the Schalkwijk and Kailath scheme for feedback communication. We then proceeded to use the regression coefficients generated by this model directly. This could be construed as erroneous: we ought to simulate the data generation, and compute the regression coefficients from the generated data. This would better model the actions of the neuroscientist who seeks to perform a Granger Causal analysis.

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Is the direction of Granger causal influence the same as the direction of information flow?}
\author{Praveen Venkatesh and Pulkit Grover\\Carnegie Mellon University}

@@

In order to determine the direction of causal influence, the Granger Causality metric (ratio of residual variances) from $\{X\}$ to $\{Y\}$ is \emph{compared} to that from $\{Y\}$ to $\{X\}$. The direction of causal influence is taken to be the direction with a greater Granger Causality metric. Further, this direction of causal influence is interpreted to be the direction of information flow -- the interpretation we question in this paper. We also note that many of these works use a spectral version of Granger Causality, that supplies this metric as a function of frequency. It then becomes possible to also determine the brain wave frequency at which these influences occur. However, we restrict our analysis to the simpler non-spectral version of Granger Causality, since it is sufficient to for the purpose of our arguments.

@@

\begin{enumerate}
	\item First, Granger Causality suffers from what we call the ``hidden node problem''. If two observed nodes receive causal influences from a third, latent node, then a causal influence may be detected between the observed nodes, even if they are independent of each other~\cite{pearl2009causality-hidden-node}. All nodes need to be observed, therefore, to avoid finding spurious influences.
	\item Second, if the measurements from each node are differentially affected by noise, then the predicted direction of causal influence might be opposite to the true direction (\cite{nalatore2007mitigating,andersson2005testing}). Measurements need to be relatively noiseless and precise in order to obtain the correct direction of causal influence.
	\item Third, subsampling the processes can produce misleading Granger Causal relations~\cite{gong2015discovering}. Performing Granger Causal analysis on subsampled time series can lead one to miss the causal influence. If pre-processing involves subsampling, then this should be done with care.
\end{enumerate}

@@

The Granger Causality and Directed Information metrics are a function of the regression coefficients estimated from the data by fitting the models given by equations~\ref{eq:ar-model-x-x} and~\ref{eq:ar-model-x-x-theta}. In our analysis, we described the data-generation model: the algorithm inspired by the Schalkwijk and Kailath scheme for feedback communication. We then proceeded to use the regression coefficients generated by this model directly. This could be construed as erroneous: we ought to simulate the data generation, and compute the regression coefficients from the generated data. This would better model the actions of the neuroscientist who seeks to perform a Granger Causal analysis.
